\section{Known Bugs}
\label{bugs}

\subsection*{Various problems related to GPLIGCfunctions.pm}
Leftover files from old installations (previous versions) in combination with an old \texttt{.gpligcrc}
file may cause gpligc to pick up an incompatible (old) \texttt{GPLIGCfunctions.pm}.
For details see \url{https://sourceforge.net/p/gpligc/bugs/2/}

\subsection*{Menu in OGIE (freeglut)}
With older freeglut versions (2.4.0) some problems arise due to the positioning and the size of the
menu, which sometimes is placed in a way, that parts of it don't fit the screen.
In this case it's possible to release the mouse pointer (key m --- also see next bug, sorry) and right-click in a different position. That way it's possible to reach all entries from the menu.
In freeglut 2.6.0 this seems to be solved.

\subsection*{Crash on activating the mousepointer (freeglut 2.4.0)}
Ogie will crash, if the mousepointer is activated by key m. This is a known bug in freeglut.
There is a patch for this in freeglut-cvs. On Gentoo systems use freeglut-2.4.0-r1 or later.

\subsection*{Menu does not work in OGIE (freeglut 2.8.0)}
Due to a bug in freeglut 2.8.0 the menu does not work at all.
Please update to freeglut 2.8.1 or later.

\subsection*{Weird speedogram plot}
If your (very old) perl-distribution provides the module Math::Complex earlier
than version 1.26, some errors will occur. Most likely negative values in
speedogram. This is a bug in the old Math::Complex perl module.

\subsection*{Annoying open dialog}
The \menu{File>Open} dialog starts always in the root folder (/). This is  a bug in the
Tk module (before 800.024). To avoid that, update to Perl/Tk module Tk-800.024 or later.

