\section{Installation}

\subsection{General Linux and Unix installation procedure}
\label{unix_install}
\label{linux_install}

This applies to all Linux and Unix operating systems.

\begin{enumerate}

\item Extract the archive: \\
\texttt{tar xvzf gpligc-version.tar.gz} \\
Change to the just created directory: \\
\texttt{cd gpligc-version}

\item Configure and build the software:\\
\texttt{./configure}\\
for options and details on configuring the build see README and the output of \texttt{./configure --help}

\item Build the software:\\
\texttt{make}

\item  \label{root}
Become root or run the next command using sudo.\\
\texttt{make install}

\item copy the example configuration file
\texttt{.ogierc} (PREFIX/share/gpligc/) to your HOME directory
and edit it according to your needs (see section~\ref{config}).

\item Make sure that Gnuplot \cite{gnuplot} is installed and in the path.
GPLIGC will also work without Gnuplot,  but you will not be able to use the plotting features.

\item Make sure that the Perl/Tk \cite{perltk} module is installed

\item Read the documentation to learn how to use gpligc \& ogie

\end{enumerate}


\subsubsection{Dependencies on ubuntu}
As I get many questions from ubuntu users because of missing dependencies, here is a list of
packages, which may need to be installed (maybe some more experienced ubuntu user can verify this list?).

\texttt{libjpeg-dev}, \texttt{freeglut3-dev},
\texttt{libgl1-mesa-dev},
\texttt{libglu1-mesa-dev},
\texttt{libosmesa-dev},
\texttt{libgps-dev},
\texttt{libqgpsmm-dev},
\texttt{gpsd},
\texttt{perl-tk},
\texttt{gnuplot (gnuplot-x11, gnuplot-qt)}


%\input{mcinstall.tex}

\input{winstall.tex}


\subsection{Additional Perl modules}
For best experience with gpligc you should install the following Perl modules:\\

\begin{itemize}
 \item \texttt{Image::ExifTool} \quad needed for photo-locator and geo-tagging. See \cite{exiftool}.
 \item \texttt{Imager} \quad needed for maps. See \cite{imager}.
\end{itemize}

there are (at least) two ways of installing Perl modules

\subsubsection{manually} You should go to the CPAN \cite{cpan} and search for the modules, download and install them.
 After downloading the archive(s), it takes the usual three commands: \\
\texttt{perl Makefile.PL} \\
\texttt{make} \\
\texttt{make install}   (as root)\\


\subsubsection{using the CPAN.pm module} If the cpan module isn't configured yet, this can be done interactively or even automated during this process. \\
\texttt{perl -MCPAN -e shell} \\
then enter \\
\texttt{install Image::ExifTool}\\
at the cpan prompt.
